---
title: "boot_uebung8_team1"
author: "Sven Bergmann, Malte Lennartz"
date: "`r Sys.Date()`"
output:
html_document:
df_print: paged
---

### Exercise 6.1.
Repeat exercise 5.3 but now with the resampling scheme given under 6.1 and compare the results.

#### Exercise 5.3.
Use the model
$$Y_i = x_i \beta + \epsilon_i, x_i = \frac{i}{n},$$
where $\epsilon_1 = x_1\delta_1, \ldots, \epsilon_n = x_n \delta_n$ and $\delta_1, \ldots, \delta_n \sim \mathcal{N}(0,\sigma^2)$ are iid., $\beta = 0.5$ and $\sigma^2 = 4$.
Note, in this case the error terms in the model, i.e. $\epsilon_i$, are not homoscedastic anymore!
Repeat the simulation studies of Exercises 5.1 and 5.2 with this model.
```{r}
n <- 100
sigma2 <- 4
beta <- 0.5
X <- seq(1, n, 1) / n
eps <- rnorm(n, 0, sqrt(sigma2))
Y <- beta * X + eps
```
```{r}
ci_boot <- function(X, Y, alpha = 0.1, m = 1000) {
  n <- length(Y)
  model <- lm(Y ~ X - 1)
  beta_hat <- model$coefficients

  X_boot <- replicate(n = m, expr = sample(x = X, size = n, replace = TRUE))

  Y_boot <- X_boot * beta_hat

  beta_hat_boot <- apply(Y_boot, 2, function(y) lm(y ~ X - 1)$coefficients)

  # epsilon_boot <- Y_boot - t(X_boot) * beta

  Sigma <- (1 / n) * t(X) %*% X

  # Wir wissen nicht, wie wir M berechnen sollen, daher die Einheitsmatrix.
  M <- diag(length(Sigma))

  t_boot <- sqrt(n) * (beta_hat_boot - beta_hat) /
    sqrt(solve(Sigma) %*% M %*% solve(Sigma))

  quantiles <- quantile(t_boot, c(alpha / 2, 1 - alpha / 2))
  right_boot <- beta_hat - sqrt(solve(Sigma) %*% M %*% solve(Sigma) * (1 / n)) * quantiles[1]
  left_boot <- beta_hat - sqrt(solve(Sigma) %*% M %*% solve(Sigma) * (1 / n)) * quantiles[2]
  return(c(ci.left.boot = left_boot, ci.right.boot = right_boot))
}
```
```{r}
suppressWarnings({
  cis <- ci_boot(X, Y, alpha = 0.1, m = 100)
  sprintf("[%.3f, %.3f]", cis[1], cis[2])
})
```
```{r}
ci <- function(X, Y, alpha) {
  n <- length(Y)
  model <- lm(Y ~ X - 1)
  beta_hat <- model$coefficients
  V <- (1 / n * (t(X) %*% X))
  res <- model$residuals
  sigma2_hat <- var(res)
  right <- beta_hat - sqrt(solve(V) * sigma2_hat / n) * qnorm(alpha / 2)
  left <- beta_hat - sqrt(solve(V) * sigma2_hat / n) * qnorm(1 - alpha / 2)
  return(c(ci.left = left, ci.right = right))
}
```
```{r}
simulate_ci <- function(X, beta, sigma2, alpha, m) {
  n <- length(X)
  eps <- rnorm(n, 0, sqrt(sigma2))
  Y <- beta * X + eps
  cis <- ci(X, Y, alpha)
  cis_boot <- ci_boot(X, Y, alpha, m)
  return(c(cis, cis_boot))
}
```
```{r}
suppressWarnings({
  set.seed(123)
  rep <- 100
  alpha <- 0.1
  m <- 1000
  sim <- replicate(rep, simulate_ci(X, beta, sigma2, alpha, m))
  ci_mean <- mean(sim[2,] - sim[1,])
  ci_mean_boot <- mean(sim[4,] - sim[3,])
  p_ci <- mean(sim[1,] < sim[3,] & sim[2,] > sim[4,])
  p_beta <- mean(sim[1,] < beta & beta < sim[2,])
  p_beta_boot <- mean(sim[3,] < beta & beta < sim[4,])
  cat('ci_mean', ci_mean, '\n')
  cat('ci_mean_boot', ci_mean_boot, '\n')
  cat('p_ci', p_ci, '\n')
  cat('p_beta', p_beta, '\n')
  cat('p_beta_boot', p_beta_boot, '\n')
})
```

### Exercise 6.2.
Use the model $$Y_i = X_i + \epsilon_i$$,
where $X_1, \ldots, X_n \sim WEIB(0.4, 4)$ are iid., $\epsilon_i = X_i\cdot\delta_i, \delta_1, \ldots, \delta_n \sim\mathcal{N}(0,4)$ are iid., $\beta = 0.5$, and $n = 50$. Repeat the simulation study of exercise 5.3 under this model
with respect to the resampling scheme given under 6.1. (To generate $WEIB(0.4, 4)$
compare Exercise 2.3).
```{r}
simulate_ci_weibull <- function(beta = 0.5, sigma2 = 4, alpha = 0.05, m = 1000) {
  X <- rweibull(n = 50, shape = 0.4, scale = 4)
  n <- length(X)
  eps <- X * rnorm(n = 50, mean = 0, sd = sqrt(sigma2))
  Y <- X * beta + eps
  cis <- ci(X, Y, alpha)
  cis_boot <- ci_boot(X, Y, alpha, m)
  return(c(cis, cis_boot))
}
```
```{r}
suppressWarnings({
  set.seed(123)
  rep <- 100
  alpha <- 0.1
  m <- 1000
  sim <- replicate(rep, simulate_ci_weibull())
  ci_mean <- mean(sim[2,] - sim[1,])
  ci_mean_boot <- mean(sim[4,] - sim[3,])
  p_ci <- mean(sim[1,] < sim[3,] & sim[2,] > sim[4,])
  p_beta <- mean(sim[1,] < beta & beta < sim[2,])
  p_beta_boot <- mean(sim[3,] < beta & beta < sim[4,])
  cat('ci_mean', ci_mean, '\n')
  cat('ci_mean_boot', ci_mean_boot, '\n')
  cat('p_ci', p_ci, '\n')
  cat('p_beta', p_beta, '\n')
  cat('p_beta_boot', p_beta_boot, '\n')
})
```
